### 2018-05-09 Final Paper/paper.tex
Cleaned up Final Paper
Made a bunch of additions. Looks good now.

\usepackage{graphicx}
\usepackage{caption}
%\usepackage{subfig}
\usepackage{subcaption}
\usepackage{listings}

@@

\section{The Need for Smoothing}
\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/unsmoothed_data"}.png}
	\caption{Look at this mess. Look upon it, and weep.}
	\label{fig:unsmoothed_data}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/grid"}.jpg}
	\caption{Hamming and Flattop Window Test: Using MATLAB code, this shows us one of the sets of data, and shows the effect of using the Hamming and Flattop window functions to dampen it. The bottom result is supposed to show us the periods detected after the application of the algorithm, however, it is not detecting the periods properly, likely due to the cancerous data. This is further confirmed by the period being detected, which is much larger than it should be, as the periods should be 20-30h at best. The explicit use versus the ones matrix indicates if there was an additional matrix used in mitigating the effects of the algorithm on the signal. Explicit use meant no additional matrix was used, ones matrix indicates that ones matrix was used.}
	\label{fig:grid}
\end{figure}

@@

\section{Zielinski et al. - Multiple Cos Fitting}
Given that this approach is the tried and true approach to period-finding, we decided to use it as a starting point for this project. Zielinski, Moore, et al. published a paper detailing various methods of period finding for circadian data, from which we took our first function, which is called mFourFit. This function (seen in Equation \ref{eq:mFourFit}) makes the assumption that any periodic data can be represented as a linear combination of cosines with different periods. It simply adds together $j$ cosines, where cosine $i$ has period $\tau/i$ and amplitude $A_i$. It also makes the assumption that it is possible to model the data using only periods that are multiples of each other (so there is one base period, $\tau$, and then the rest of the periods are $\tau$ divided by an integer, $i$). An example of a function generated using mFourFit and a random set of parameters can be seen in Figure \ref{fig:mFourFit_ex}. 
Zielinski also described another function, FFT-NLLS, which can be seen in Equation \ref{eq:FFT-NLLS}. FFT-NLLS is very similar to mFourFit, but it doesn’t assume that the periods are multiples of each other. Instead, it uses $j$ individual periods that are all found during the parameter-finding phase. This potentially provides a better fit to messier data, but at the same time it dramatically increases the computation cost of parameter-fitting.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/FourFit_GA_match"}.png}
	\caption{A plot of artificially generated data that has then been fit to using the mFourFit equation.}
	\label{fig:mFourFit_ex}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/NLLS_GA_realdata"}.png}
	\caption{A plot of our real data, and the fit we found using a genetic algorithm with FFT-NLLS. The fit does not get the individual oscillations, but it does find the general trend of the data.}
	\label{fig:FFT-NLLS_real}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/corrected_poly"}.png}
	\caption{Using the third-degree polynomial fit to find the downwards trend of our data, and correcting the data based on that downwards trend.}
	\label{fig:correcting_with_poly}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/first_drv_fail"}.png}
	\caption{Using the first derivative to find peaks.}
	\label{fig:first_drv_fail}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/third_drv_fail"}.png}
	\caption{Using the third derivative to find peaks.}
	\label{fig:third_drv_fail}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/pv_det_final"}.png}
	\caption{Our final peak and valley detection process. This process worked consistently for every dataset that we tried it on.}
	\label{fig:pv_det_final}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{{"plots/corrected_poly"}.png}
	\caption{A plot of the found period over time of one of our datasets. We also plotted the third-degree polynomial fit of the found period data in order to get a continuous function of the period over time.}
	\label{fig:per_over_time}
\end{figure}

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}
\bibitem{FourFit}
Zielinski T, Moore AM, Troup E, Halliday KJ, Millar AJ (2014). \emph{Strengths and Limitations of Period Estimation Methods for Circadian Data}. PLoS ONE 9(5):e96462. doi:10.1371/journal.pone.0096462

@Misc{,
  author =    {Eric Jones and Travis Oliphant and Pearu Peterson and others},
  title =     {{SciPy}: Open source scientific tools for {Python}},
  year =      {2001--},
  url = "http://www.scipy.org/",
  note = {[Online; accessed <today>]}
}

@Article{Hunter:2007,
  Author    = {Hunter, J. D.},
  Title     = {Matplotlib: A 2D graphics environment},
  Journal   = {Computing In Science \& Engineering},
  Volume    = {9},
  Number    = {3},
  Pages     = {90--95},
  abstract  = {Matplotlib is a 2D graphics package used for Python
  for application development, interactive scripting, and
  publication-quality image generation across user
  interfaces and operating systems.},
  publisher = {IEEE COMPUTER SOC},
  doi       = {10.1109/MCSE.2007.55},
  year      = 2007
}

@misc(MatlabOTB,
  key = {MATLAB Optimization Toolbox},
  title = {MATLAB Optimization Toolbox},
  year = {2017}
  note = {The MathWorks, Natick, MA, USA}
}

>>>>

\usepackage{graphicx}
\usepackage{caption}
%\usepackage{subfig}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{cite}

@@

\section{The Need for Smoothing}
\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/unsmoothed_data"}.png}
	\caption{Look at this mess. Look upon it, and weep.}
	\label{fig:unsmoothed_data}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/grid"}.jpg}
	\caption{Hamming and Flattop Window Test: Using MATLAB code, this shows us one of the sets of data, and shows the effect of using the Hamming and Flattop window functions to dampen it. The bottom result is supposed to show us the periods detected after the application of the algorithm, however, it is not detecting the periods properly, likely due to the cancerous data. This is further confirmed by the period being detected, which is much larger than it should be, as the periods should be 20-30h at best. The explicit use versus the ones matrix indicates if there was an additional matrix used in mitigating the effects of the algorithm on the signal. Explicit use meant no additional matrix was used, ones matrix indicates that ones matrix was used.}
	\label{fig:grid}
\end{figure}

@@

\section{Zielinski et al. - Multiple Cos Fitting}
Given that this approach is the tried and true approach to period-finding, we decided to use it as a starting point for this project. Zielinski, Moore, et al. published a paper detailing various methods of period finding for circadian data \cite{Zielinski}, from which we took our first function, which is called mFourFit. This function (seen in Equation \ref{eq:mFourFit}) makes the assumption that any periodic data can be represented as a linear combination of cosines with different periods. It simply adds together $j$ cosines, where cosine $i$ has period $\tau/i$ and amplitude $A_i$. It also makes the assumption that it is possible to model the data using only periods that are multiples of each other (so there is one base period, $\tau$, and then the rest of the periods are $\tau$ divided by an integer, $i$). An example of a function generated using mFourFit and a random set of parameters can be seen in Figure \ref{fig:mFourFit_ex}. 
Zielinski also described another function, FFT-NLLS, which can be seen in Equation \ref{eq:FFT-NLLS}. FFT-NLLS is very similar to mFourFit, but it doesn’t assume that the periods are multiples of each other. Instead, it uses $j$ individual periods that are all found during the parameter-finding phase. This potentially provides a better fit to messier data, but at the same time it dramatically increases the computation cost of parameter-fitting.

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/FourFit_GA_match"}.png}
	\caption{A plot of artificially generated data that has then been fit to using the mFourFit equation.}
	\label{fig:mFourFit_ex}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/NLLS_GA_realdata"}.png}
	\caption{A plot of our real data, and the fit we found using a genetic algorithm with FFT-NLLS. The fit does not get the individual oscillations, but it does find the general trend of the data.}
	\label{fig:FFT-NLLS_real}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/corrected_poly"}.png}
	\caption{Using the third-degree polynomial fit to find the downwards trend of our data, and correcting the data based on that downwards trend.}
	\label{fig:correcting_with_poly}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/first_drv_fail"}.png}
	\caption{Using the first derivative to find peaks.}
	\label{fig:first_drv_fail}
\end{figure}

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/third_drv_fail"}.png}
	\caption{Using the third derivative to find peaks.}
	\label{fig:third_drv_fail}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/pv_det_final"}.png}
	\caption{Our final peak and valley detection process. This process worked consistently for every dataset that we tried it on.}
	\label{fig:pv_det_final}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{{"plots/corrected_poly"}.png}
	\caption{A plot of the found period over time of one of our datasets. We also plotted the third-degree polynomial fit of the found period data in order to get a continuous function of the period over time.}
	\label{fig:per_over_time}
\end{figure}

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{3}
\bibitem{Zielinski}
Zielinski T, Moore AM, Troup E, Halliday KJ, Millar AJ (2014). \emph{Strengths and Limitations of Period Estimation Methods for Circadian Data}. PLoS ONE 9(5):e96462. doi:10.1371/journal.pone.0096462

\bibitem{SciPy}
Eric Jones and Travis Oliphant and Pearu Peterson and others (2001--). \emph{{SciPy}: Open source scientific tools for {Python}}. "http://www.scipy.org/", {[Online; accessed 5/8/2018]}

\bibitem{Matplotlib}
Hunter, J. D. (2007). \emph{Matplotlib: A 2D graphics environment}. Computing In Science \& Engineering, IEEE COMPUTER SOC, 10.1109/MCSE.2007.55.

\bibitem{MatLab}
(2017). \emph{MATLAB Optimization Toolbox}. The MathWorks, Natick, MA, USA.
